---
title: "Homework 6"
format: html
editor: visual
---

## Task 1: Conceptual Questions

1.  What is the purpose of the lapply() function? What is the equivalent purrr function?

> lapply() applies a single function to every element in a list. The purrr equivalent is map().

2.  Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall") on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify method = "kendall" when calling lapply())

> You would add method = 'kendall' as an additional argument in lapply(). It would be lapply(x, Fun = cor, method = 'kendall').

3.  What are two advantages of using purrr functions instead of the BaseR apply family?

> purrr is more consistent and has additional helper functions.

4.  What is a side-effect function?

> A side-effect function does something other than the 'main' goal of the function. There is additional action taken by r.

5.  Why can you name a variable sd in a function and not cause any issues with the sd function?

> A custom function has its own temporary environment. Anything created inside the function only lives there.

## Task 2: Writing R Functions

Below, a function was created to calculate the root mean square error (RMSE). The function takes in a vector of observations that must be all be numeric and a vector of predictions that must all be numeric. Those two vectors must also be equal in length (each observation having a corresponding prediction). The function includes an optional argument to handle any missing values in a vector. The function can't handle missing values by default, so an argument must be provided if a missing value is present.

```{r}
getRMSE<-function(obs, pres, ...){
if(length(obs)!=length(pres)){
  stop('Different number of responses and predictions.')
}
z<- obs - pres
vn<-z^2
g<-mean(vn, ...)
x<-sqrt(g)
return(x)
}
```

Testing our function:

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

getRMSE(resp, pred)
```

Things change if a missing value is added to a vector.

```{r}
resp[1]<-NA
resp[6]<-NA

getRMSE(resp, pred)
```

The function cannot run with the missing value. We must specify that we want missing values removed.

```{r}
getRMSE(resp, pred, na.rm=TRUE)
```

Below, a function was created to calculate the mean absolute deviation (MAE). The function takes in a vector of observations that must be all be numeric and a vector of predictions that must all be numeric. Those two vectors must also be equal in length (each observation having a corresponding prediction). The function includes an optional argument to handle any missing values in a vector. The function can't handle missing values by default, so an argument must be provided if a missing value is present.

```{r}
getMAE<-function(obs, pres, ...){
  if(length(obs)!=length(pres)){
  stop('Different number of responses and predictions.')
}
z<- obs - pres
vn<-abs(z)
g<-mean(vn, ...)
return(g)
}
```

Testing our function:

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

getMAE(resp, pred)
```

Things change if a missing value is added to a vector.

```{r}
resp[2]<-NA
resp[7]<-NA

getRMSE(resp, pred)
```

The function cannot run with the missing value. We must specify that we want missing values removed.

```{r}
getMAE(resp, pred, na.rm=TRUE)
```

Now we will build a function that can call either of the above functions or both, with both being returned by default.

```{r}
getMEAS<-function(obs, pres, RMSE=TRUE, MAE=TRUE, ...) {
  if (!is.vector(obs)|!is.atomic(obs)|!is.numeric(obs)){
    stop('The response vector does not meet function requirements.')
  }
  if (!is.vector(pres)|!is.atomic(pres)|!is.numeric(pres)){
    stop('The prediction vector does not meet function requirements.')
  }
    if(length(obs)!=length(pres)){
  stop('Different number of responses and predictions.')
}
  if (RMSE==TRUE & MAE==FALSE) {
    x<-getRMSE(obs, pres, ...)
    my_list<-list(RMSE=x)
  }
  if (RMSE==FALSE & MAE==TRUE){
    y<-getMAE(obs, pres, ...)
    my_list<-list(MAE=y)
  }
  if (RMSE==TRUE & MAE==TRUE){
    x<-getRMSE(obs, pres, ...)
    y<-getMAE(obs, pres, ...)
    my_list<-list(RMSE=x, MAE=y)
  }
return(my_list)
}
```

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

getMEAS(resp, pred)

```

The function can also call just one of the measurements.

```{r}
getMEAS(resp, pred, RMSE=FALSE)
```

Let's give our function vectors with missing values.

```{r}
resp[3]<-NA
resp[10]<-NA

getMEAS(resp, pred)
```

We can also specify removing missing values for this function.

```{r}
getMEAS(resp, pred, na.rm=TRUE)
```

The function will not work if the inputs are not of the correct type. For example, the function will not work if the prediction values are stored in a data frame, not a vector. I was unable to actually run this example. It correctly produced the error message, but then I was unable to render the website because of the error. I was unable to find my way around it.

```{r}
#| error: true
pred<-data.frame(pred)

getMEAS(resp, pred, na.rm=TRUE)
```

## Task 3: Querying an API and a Tidy-Style Function

```{r}
#| include: false
#| echo: false
library(httr)
library(jsonlite)
library(purrr)
library(lubridate)
library(dplyr)
```

Now we are going to create a function that queries newsapi.org. Our function gives the user the option to specify a subject they are interested, the start date to search from, and an API key. Then the function will call data from news articles that meet those specifications.

The function uses `get()` to access the API. The function then parses that data for actual article information ('content'). Then we use `pluck()` to grab the 'articles' element and that is returned as a data frame.

```{r}
get_NEWS_API<-function(topic, startdate, APIkey){
  start<-'https://newsapi.org/v2/everything?q='
  topic_url<-topic
  continue<-'&from='
  startdate_url<-startdate
  continue2<-'&sortBy=popularity&apiKey='
  APIkey_url<-APIkey
  try<-paste0(start,topic_url,continue,startdate_url,continue2,APIkey_url)
  try2<-httr::GET(try)
  try3<-fromJSON(rawToChar(try2$content))
  try4<-purrr::pluck(try3,"articles")
  return(try4)
}
```

We'll use our function to create two objects, one about football and the other about baseball.

```{r}
football<-get_NEWS_API('football','2024-10-14','42b42fbb54a644faa2b984b30dbe20ae')

baseball<-get_NEWS_API('baseball','2024-10-14','42b42fbb54a644faa2b984b30dbe20ae')
```

We'll summarize the names of the sources from each object. The tables provide counts for each source name.

```{r}
table(football$source$name)
```

```{r}
table(baseball$source$name)
```

Next, we'll convert `publishedAt` to a date. Then we'll sort the data by their published date/time, ascending. Then we'll use `lag()` to grab `publishedAt` from the previous observation and create `prev_pub`. We can subtract `prev_pub` from `publishedAt` to calculate the time in seconds since the last article was published. `pub_diff` shows that time in seconds.

```{r}
football<-football |>
  mutate(publishedAt=ymd_hms(publishedAt)) |>
  arrange(publishedAt) |>
  mutate(prev_pub=lag(publishedAt), pub_diff = as.numeric(publishedAt - prev_pub))

baseball<-baseball |>
  mutate(publishedAt=ymd_hms(publishedAt)) |>
  arrange(publishedAt) |>
  mutate(prev_pub=lag(publishedAt), pub_diff = as.numeric(publishedAt - prev_pub))
```

Now we'll use our `football` data frame and summarize the `publishedAt` and `pub_diff` variables.

```{r}
small_football<-football |>
  select(publishedAt, pub_diff)

map(small_football, \(x) c(mean(x, na.rm=TRUE), median(x, na.rm=TRUE), sd(x, na.rm=TRUE)))
```
